Graph Neural Networks For Multi-Image Matching

### 2019-11-13 latex/iclr2020/main.tex
More ICLR revisions

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D feature positions from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \citep{lowe2004distinctive} combined with RANSAC \citep{fischler1981random} have been the standard for decades.
However RANSAC-based approaches are limited to matching pairs of images, which can lead to global inconsistencies in the matching.
% TODO: Mention trifocal tensors
Other works, such as \cite{wang2017multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent.

@@

In this work, we address these concerns using Graph Neural Networks (GNNs).
The proposed method works directly on the graph of correspondences between the image features, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss, the cycle consistency loss, to train the network, thus avoiding the difficulty of expensive hand-labeling.
We use geometric consistency losses to aid training, though no geometric information is used at inference time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss \citep{zhou2015multi, leonardos2016distributed}.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster to achieve comparable accuracy than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K dataset \citep{li2010location} to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised multi-view cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section \ref{sec:cycconsist}).
\cite{pachauri2013solving} and \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the assumption of low Gaussian noise is not realistic.
\cite{zhou2015multi} and \cite{wang2017multi} use more sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
\cite{leonardos2016distributed} implement a distributed optimization scheme to solve for cycle consistency.
\textcolor{red}{\cite{swoboda2019convex} implement a convex relaxation of the low dimensional embedding problem.}
\textcolor{red}{\cite{shi2016tensor} use tensor power iterations to solve the matching problem, also taking into account the intra-image matching graph.}
As an alternative to optimization based techniques, \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
Moving away from feature matching, \cite{zach2010disambiguating} uses cycle-consistency-like constraints on pose graphs quite effectively.
To the best of our knowledge, we are the first to use graph neural networks for multi-image matching.

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets such as \cite{zagoruyko2015learning}; \cite{yi2016lift}; and \cite{brachmann2017dsac}.
This approach is limited if one does not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as \cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, \cite{yi2018learning} attempts to improve correspondences by learning match probabilities for two-view RANSAC for greater robustness and speed.
Like us \cite{zhu2017unpaired} use cycle consistency in their loss; however their method is for image generation and is restricted to pairwise cycle consistency.
Our method can be applied to 3 or more images. 
\cite{hartmann2017learned} learns multi-image matching but requires heavy supervision from 3D object reconstructions, which can be difficult or expensive to obtain.
\textcolor{red}{\cite{zhou2015flowweb}, unlike our method, uses dense correspondence, but uses cycle consistency to find semantic matches accross multiple views.}
\textcolor{red}{\cite{suwajanakorn2018discovery} find 3 dimensional latent keypoints, trained using ground truth rotation and translation. However, their method is restricted to a limited number of object categories, which is different from the SfM setting we are considering here.}

@@

Let $\mM$ be the noiseless set of matches between our features, with $\mM_{ij}$ being the partial permutation representing the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then for all $i, j, k$:
\begin{equation}
\mM_{ij} = \mM_{ik} \mM_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mX_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mM_{ij} = \mX_{i}\mX_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth correspondences but using unsupervised methods and geometric side losses.
Thus we test against ground truth correspondence graph adjacency matrices computed from the bundle adjustment output.
Our method performs better than 35 iteration of the MatchALS \citep{zhou2015multi} method, but does not perform as well as 50 iterations.
% We do not perform as well as the Projected Gradient Descent Doubly Stochastic (PGDDS) \citep{leonardos2016distributed} but we perform significantly faster than them.
We perform better than a simple eigenvalue based method \citep{pachauri2013solving}.
Note that we perform much better in $L_1$ performance rather than $L_2$, as we optimized the network weights using an $L_1$ loss.
}
\label{tab:rome16ktab}
\end{table*}

@@

We compare our method to spectral and optimization based baselines with different maximum iteration cutoffs.
Figure \ref{fig:errorlines} illustrates this by plotting the means of various metrics and their $25^{th}$ and $75^{th}$ percentiles, with table \ref{tab:rome16ktab} giving the exact numbers.
Our network, though only using 8 message passes, has comparable accuracy to MatchALS \citep{zhou2015multi} run 35 to 45 iterations, with an equivalent message passing step at each phase.
Although our method does not outperform the Projected Gradient Descent - Doubly Stochastic (PGDDS) \citep{leonardos2016distributed} method, we perform comparably to them in the ROC AUC metric.

\subsubsection{Graffiti Dataset} \label{sec:graffiti}
We run our trained model on the more Graffiti Dataset from the Affine Covariant Regions dataset (formatted to be able to be input to our model properly).
The Graffiti Dataset is the most common benchmark used in feature matching algorithms (e.g. \cite{leonardos2016distributed, zhou2015multi}).
The results are shown in figure \ref{fig:errorlines} in the rightmost figure.
As the graffiti dataset is very small (only 6 views total), we were not able to train on it.
We randomly permute the intra-image order of the features to add some variance - by design the GNN outputs the same result each time, while the optimization methods have a very small amount of variance.
The transferred results of Graffiti are similar to the test error of Rome16K - smaller $L_1$ error and comparable ROC error.
This shows that the GNNs trained in Rome16K generalize similarly to the optimization based methods.

>>>>

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D feature positions from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \citep{lowe2004distinctive} combined with RANSAC \citep{fischler1981random} have been the standard for decades.
However RANSAC-based approaches are limited to matching pairs of images, which can lead to global inconsistencies in the matching.
% TODO: Mention trifocal tensors
Other works, such as \cite{wang2018multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent.

@@

In this work, we address these concerns using Graph Neural Networks (GNNs).
The proposed method works directly on the graph of correspondences between the image features, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss, the cycle consistency loss, to train the network, thus avoiding the difficulty of expensive hand-labeling.
We use geometric consistency losses to aid training, though no geometric information is used at inference time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss \citep{zhou2015multi, leonardos2017distributed}.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster to achieve comparable accuracy than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K dataset \citep{li2010location} to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised multi-view cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section \ref{sec:cycconsist}).
\cite{pachauri2013solving} and \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the assumption of low Gaussian noise is not realistic.
\cite{zhou2015multi} and \cite{wang2017multi} use more sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
\cite{leonardos2017distributed} implement a distributed optimization scheme to solve for cycle consistency.
\textcolor{red}{\cite{swoboda2019convex} implement a convex relaxation of the low dimensional embedding problem.}
\textcolor{red}{\cite{shi2016tensor} use tensor power iterations to solve the matching problem, also taking into account the intra-image matching graph.}
As an alternative to optimization based techniques, \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
Moving away from feature matching, \cite{zach2010disambiguating} uses cycle-consistency-like constraints on pose graphs quite effectively.
To the best of our knowledge, we are the first to use graph neural networks for multi-image matching.

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets such as \cite{zagoruyko2015learning}; \cite{yi2016lift}; and \cite{brachmann2017dsac}.
\textcolor{red}{\cite{rocco2018neighbourhood} use a dense matching tensor with a weakly supervised loss to learn good features for matching. \cite{zhou2019learn} has shown that these learned representations have limited ability for improving SfM tasks.}
% This approach is limited if one does not have the ability to get the ground truth correspondences.

There are other methods to build correspondences such as \cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, \cite{yi2018learning} attempts to improve correspondences by learning match probabilities for two-view RANSAC for greater robustness and speed.
Like us \cite{zhu2017unpaired} use cycle consistency in their loss; however their method is for image generation and is restricted to pairwise cycle consistency.
Our method can be applied to 3 or more images. 
\cite{hartmann2017learned} learns multi-image matching but requires heavy supervision from 3D object reconstructions, which can be difficult or expensive to obtain.
\textcolor{red}{\cite{zhou2015flowweb}, unlike our method, uses dense correspondence, but uses cycle consistency to find semantic matches accross multiple views.}
\textcolor{red}{\cite{suwajanakorn2018discovery} find 3 dimensional latent keypoints, trained using ground truth rotation and translation. However, their method is restricted to a limited number of object categories, which is different from the SfM setting we are considering here.}

@@

Let $\mM$ be the noiseless set of matches between our features, with $\mM_{ij}$ being the partial permutation representing the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then for all $i, j, k$:
\begin{equation}
\mM_{ij} = \mM_{ik} \mM_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2017distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mX_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mM_{ij} = \mX_{i}\mX_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth correspondences but using unsupervised methods and geometric side losses.
Thus we test against ground truth correspondence graph adjacency matrices computed from the bundle adjustment output.
Our method performs better than 35 iteration of the MatchALS \citep{zhou2015multi} method, but does not perform as well as 50 iterations.
% We do not perform as well as the Projected Gradient Descent Doubly Stochastic (PGDDS) \citep{leonardos2017distributed} but we perform significantly faster than them.
We perform better than a simple eigenvalue based method \citep{pachauri2013solving}.
Note that we perform much better in $L_1$ performance rather than $L_2$, as we optimized the network weights using an $L_1$ loss.
}
\label{tab:rome16ktab}
\end{table*}

@@

We compare our method to spectral and optimization based baselines with different maximum iteration cutoffs.
Figure \ref{fig:errorlines} illustrates this by plotting the means of various metrics and their $25^{th}$ and $75^{th}$ percentiles, with table \ref{tab:rome16ktab} giving the exact numbers.
Our network, though only using 8 message passes, has comparable accuracy to MatchALS \citep{zhou2015multi} run 35 to 45 iterations, with an equivalent message passing step at each phase.
Although our method does not outperform the Projected Gradient Descent - Doubly Stochastic (PGDDS) \citep{leonardos2017distributed} method, we perform comparably to them in the ROC AUC metric.

\subsubsection{Graffiti Dataset} \label{sec:graffiti}
We run our trained model on the more Graffiti Dataset from the Affine Covariant Regions dataset (formatted to be able to be input to our model properly).
The Graffiti Dataset is the most common benchmark used in feature matching algorithms (e.g. \cite{leonardos2017distributed, zhou2015multi}).
The results are shown in figure \ref{fig:errorlines} in the rightmost figure.
As the graffiti dataset is very small (only 6 views total), we were not able to train on it.
We randomly permute the intra-image order of the features to add some variance - by design the GNN outputs the same result each time, while the optimization methods have a very small amount of variance.
The transferred results of Graffiti are similar to the test error of Rome16K - smaller $L_1$ error and comparable ROC error.
This shows that the GNNs trained in Rome16K generalize similarly to the optimization based methods.
